### 2019-12-06 reports/paper.tex
start writing

\documentclass[10pt,conference]{IEEEtran}
\usepackage[dvipsnames]{xcolor}
\usepackage{standalone}
\usepackage{tikz}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{booktabs}
\usepackage{tabu}
\usepackage[letterpaper]{geometry}
\usepackage[colorlinks]{hyperref}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2cm,rmargin=2cm,columnsep=0.8cm}
\usetikzlibrary{positioning, calc, shapes, fit, backgrounds, patterns}

\title{\textbf{Hardware accelerator topologies in heterogeneous systems: an evaluation in the
    security domain}}
\author{Bruno Morais}
\date{}

\begin{document}

\maketitle

\begin{abstract}
  Hardware accelerators (HWACCs) have driven performance gains for
  domain specific applications which leverage highly optimized computation
  cores. However, the traditional accelerator access topology as an add-on to
  general purpose processors through shared memory space/bus, reveals performance
  issues as the number of HWACCs grow and systems become
  communication-bound.
  On the other hand, efficient development and research of HWACCs
  that are closer to a general purpose processor's datapath is seen as a
  prohibitive task due to the required microarchitecture knowledge and lack of
  tools to explore the design space. This has in turn hindered the exploration of
  such architectures for gains in either performance or security spectrums.
  RISC-V projects made available in the last few years provide a framework that
  greatly eases microarchitecture research, enabling further experimentation
  into security and performance aspects for heterogeneous platforms.
  This work uses a customized RISC-V SoC to explore performance trade-offs
  between different HWACC topologies (in-datapath, memory-mapped) for embedded
  systems in the security domain by implementing an IP core for cryptographic
  primitive acceleration.
  % results also go in here

\end{abstract}

\section{Introduction}

\section{Related work}

\subsection{Hardware accelerators?}
% Tightly vs loosely coupled hwaccs

\subsection{RISC-V instruction set extensions}
% Xcrypto

\section{Background}

\subsection{Heterogeneous systems and HWACCs}

\begin{figure}
  \centering
  \include{media/loosely_acc}
  \caption{Loosely-coupled HWACCs: typical memory mapped HWACC usage in a heterogenous architecture.}
\end{figure}

\begin{figure}
  \centering
  \scalebox{0.75}{
  \begin{minipage}{0.5\textwidth}
    \centering
    \include{media/tightly_acc}
  \end{minipage}
  }
  \caption{Tightly-coupled HWACCs: accelerator is moved into CPU datapath,
    similar to Functional Units (FUs).}
\end{figure}

\begin{figure}
  \centering
  \include{media/hybrid_acc}
  \caption{Hybrid cryptographic HWACCs: \textbf{upper}: HWACC is directly accessible through CPU datapath and
    integrated with memory hierarchy; \textbf{lower}: two operation modes; I: batch
  background processing of memory region (CPU not taxed); II: cryptographic
  operation where plaintext is only visible inside CPU datapath.}
\end{figure}

\subsection{Embedded security: lightweight ciphers}

The industry standard ciphers might impose too much of a burden on deeply
embedded, low power systems that need the ability to communicate securely over
an unsecured channel. Thus, different organizations have hosted competitions to
foster the development of lightweight ciphers, targeting both resource-limited
software and hardware implementations.

In this evaluation, it is assumed that the object of study is such a system,
with strict area and power limitations. The Simon~\cite{Beaulieu2015} cipher family is chosen as a
lightweight cipher; it is a hardware-optimized cipher family developed by the
NSA. Although faced with controversies due to polarized views towards the
authors, it is an ISO standardized cipher.

\subsection{RISC-V ecosystem}

\section{Experimental setup}

\section{Evaluation}

\section{Future work}

As future work, there is much room for improvement; better understanding of how
to leverage the memory coherence hierarchy in the hybrid accelerator model
should lead to better performance.

Also, a set of different, reutilizable cryptographic primitives can be further
explored and integrated into a larger multi-accelerator framework for
acceleration of other demanded cryptographic functions such as hash functions
and possible compositions of primitives for message authentication, for instance.

These require a more thorough and complex modification of the available
toolchain and datapath microarchitecture, as it is likely that the standard RoCC
interface will not be able to provide the desired flexibility and control
granularity levels for detailed exploration.

\section{Conclusions}

\bibliography{paper}
\bibliographystyle{ieeetr}

\end{document}