### 2019-12-10 reports/presentation.tex
partial presentation

\documentclass[dvipsnames]{beamer}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{wrapfig}
\usetikzlibrary{positioning, calc, shapes, fit, backgrounds, patterns}
\addtobeamertemplate{navigation symbols}{}{%
  \usebeamerfont{footline}%
  \usebeamercolor[fg]{footline}%
  \hspace{1em}%
  \insertframenumber/\inserttotalframenumber
}
\setbeamercolor{footline}{fg=NavyBlue}
\setbeamerfont{footline}{series=\bfseries}
\title{Hardware accelerator coupling in heterogeneous systems: an evaluation in the
  security domain}
\author{Bruno S M M Morais}
\date{December 10, 2019}

\begin{document}

\maketitle

\begin{frame}{Heterogeneous systems}
  \begin{itemize}
  \item Specialized HWACCs required for energy efficient embedded / edge computing
  \item Explosion in \# of ACCs in recent SoCs, specially with deployment of
    machine learning, DSP algorithms for instance
  \end{itemize}
\end{frame}

\begin{frame}{Loosely-coupled accelerators}
  \begin{columns}[T]
    \begin{column}{0.35\textwidth}
      \resizebox{0.35\textwidth}{!}{
        \begin{minipage}{0.5\textwidth}
          \centering
          \include{media/loosely_acc}
        \end{minipage}
      }
    \end{column}
    \hfill
    \begin{column}{0.65\textwidth}
      \centering
      \begin{itemize}
      \item Access through shared memory bus
      \item Usually interrupts CPU when task is done
      \item With increase of \# of ACCs, communication becomes bottleneck
      \item CPU typically orchestrates compositions of HWACC functions
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Tightly-coupled accelerators}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item ACC moved into CPU datapath
        \item Requires ISA modification/extensions
        \item More difficult to explore, requires extensive modification of
          $\mu$Architecture and tools
      \end{itemize}
    \end{column}
    \hfill
    \begin{column}{0.5\textwidth}
      %\centering
      \resizebox{0.45\textwidth}{!}{
        \begin{minipage}{0.75\textwidth}
          \centering
          \include{media/tightly_acc}
        \end{minipage}
      }
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{RISC-V ecosystem}
  \begin{columns}[T]
    \begin{column}{0.6\textwidth}
      \begin{itemize}
      \item FOSS ISA specification and many implementations
      \item ISA is built to be extensible
      \item RoCC interface enables attaching HWACCs to datapath easily
        \begin{itemize}
        \item 4 custom opcodes for routing to 4 HWACCs
        \item Contents of two registers passed into HWACC
        \item Optionally response from HWACC written to a register
        \item Optional interface for requests into L1 cache
        \end{itemize}
      \end{itemize}
    \end{column}
    \hfill
    \begin{column}{0.4\textwidth}
      \centering
      \includegraphics[width=0.75\textwidth]{media/Tall_2.pdf}
      \includegraphics[width=1.0\textwidth]{media/rocc.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{The Simon cipher family}
  \begin{columns}[T]
    \begin{column}{0.35\textwidth}
    \include{media/simonround}
    \end{column}
    \hfill
  \begin{column}{0.65\textwidth}
    \begin{itemize}
    \item Optimized for hardware implementations
    \item Only bitwise rotate, bitwise XOR and bitwise AND
    \item 64/128 (44 rounds) and 128/128 (128 rounds) configurations
    \end{itemize}
  \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Hybrid crypto accelerator}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Batch encrypt/decrypt
      \item Transparent encrypted memory load/store
      \item Direct access into coherent memory through DMA
      \item In this work, HWACC accesses into memory at L2 cache level
      \end{itemize}
    \end{column}
    \hfill
    \begin{column}{0.5\textwidth}
      % \centering
      \resizebox{0.3\textwidth}{!}{
        \begin{minipage}{0.5\textwidth}
          \centering
          \include{media/hybrid_acc}
        \end{minipage}
      }
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Experimental setup}
  \begin{minipage}[0.2\textheight]{\textwidth}
    \begin{columns}[T]
      \begin{column}{0.8\textwidth}
        \begin{itemize}
        \item Three Simon accelerators
        \begin{itemize}
        \item MMIO (loosely-coupled)
        \item RoCC (tightly-coupled)
        \item RoCC + DMA (hybrid)
          \end{itemize}
        \end{itemize}
      \end{column}
      \begin{column}{0.2\textwidth}
        \includegraphics[width=2.5cm]{media/cogs.jpeg}
      \end{column}
    \end{columns}
  \end{minipage}
  \begin{itemize}
  \item Both in-order (Rocket) and out-of-order (BOOM) cores
  \item Performance: eight benchmarks run on cycle-accurate simulations
    \begin{itemize}
    \item Software only: ECB and CBC mode
    \item MMIO accelerator, single: ECB and CBC mode
    \item MMIO accelerator, automatic: ECB mode
    \item Tightly-coupled accelerator: ECB and CBC mode
    \item Hybrid accelerator: ECB mode
    \end{itemize}
  \item Area \& power: HWACC synthesized for Xilinx FPGA, post-implementation
    metrics extracted
  \end{itemize}
\end{frame}

\begin{frame}{Results: performance}
  \begin{columns}[T]
  \begin{column}{0.5\textwidth}
  \includegraphics[width=1.0\textwidth]{media/ecb_speedup}
  \includegraphics[width=1.0\textwidth]{media/cbc_speedup}
  \end{column}
  \begin{column}{0.5\textwidth}
  \begin{itemize}
    \item Hybrid in batch mode is fastest, followed by regular MMIO
    \item Gains with RoCC accelerator, but is slower than 2-way BOOM SW only
    \item Need to understand RoCC interface timing better
    \item CBC mode performance very similar to EBC in comparable configurations
  \end{itemize}
  \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Results: area \& power}
\end{frame}

\begin{frame}{Results: security considerations}
\end{frame}

\begin{frame}{Future work}
\end{frame}

\begin{frame}{References}
\end{frame}

\begin{frame}{Questions?}
  \centering
  \huge\bfseries ?
\end{frame}

\end{document}